BFS search in miniKanren

### 2019-03-13 Paper/paper.tex
remove timer
complete description of BFS

%%% NOTE %%%
%
% - deadline?
%
%   last year submission deadline May 15, 2018
%
%%%%%%%%%%%%

@@

The results are not cost-respecting. For example, \texttt{'(a a)} occurs before
\texttt{'(b)} while \texttt{'(a a)} is associated with a higher cost. 
The problem is that iDFS strategy prioritizes the first \texttt{conde} case
considerablely. In general, when every conde case are equally productive, the iDFS
strategy takes $1/2^{i}$ answers from the $i$-th case, except the last case,
which share the same portion as the second last.

@@

\section{change search strategy}

Now we change the search strategy and optimize the system. The whole process is completed in three steps, corresponding to 4 versions of miniKanren. The initial version, \texttt{mk-0}, is exactly the miniKanren in \emph{The Reasoned Schemer, 2nd Edition}.

@@

\texttt{mk-1} is not efficient in two aspects. \texttt{append-inf} need to copy all \texttt{cons} cells of two input stream when the first one is infinite. Besides, \texttt{mk-1} generates answers of the same cost at once, even when only a small portion is queried. We solves the two problems in the next two subsections.

\subsection{stepstone of optimization}

(eqv to Seres's BFS ?)

Make irrelevant parts in mK representation-independent w.r.t. search space, and combine mature part and immature part with cons.

\subsection{optimization}

The goal is to express BFS explicitly with queue, so that the system doesn't generate all answers of the same cost at once.

Interesting changes: (1) put thunks in a list; (2) change force-inf (introduced in 4.B) so that it can make progress in all thunks (3) use a queue to manage thunks in take-inf.

>>>>

%%% NOTE %%%
%
% - deadline?
%
%   miniKanren workshop deadline May 15, 2018
%
%%%%%%%%%%%%

@@

The results are not cost-respecting. For example, \texttt{'(a a)} occurs before
\texttt{'(b)} while \texttt{'(a a)} is associated with a higher cost. 
The problem is that iDFS strategy prioritizes the first \texttt{conde} case
considerablely. In general, when every conde case are equally productive, the iDFS
strategy takes $1/2^{i}$ answers from the $i$-th case, except the last case,
which share the same portion as the second last one.

@@

\section{breadth-first search}

In this section we change the search strategy to breadth-first search and optimize it. The whole process is completed in three steps, corresponding to 3 versions of miniKanren. We start with, \texttt{mk-0}, the miniKanren in \emph{The Reasoned Schemer, 2nd Edition}.

@@

\texttt{mk-1} is not efficient in two aspects. \texttt{append-inf} need to copy all \texttt{cons} cells of two input streams when the first stream has a non-trivial immature part. Besides, \texttt{mk-1} computes answers of the same cost at once, even when only a portion is queried. We solves the two problems in the next subsections.

\subsection{\texttt{mk-3}, optimized breadth-first search}

We avoid generating same-cost answers at once by expressing BFS with a queue. The elements of the queue are delayed computation, represented by thunks. Every \texttt{mk-1} stream has zero or one thunk, so we have no interesting way to manage it. Therefore we change the representation of immature parts from thunks to lists of thunks. As as consequence, we also change the way to combine mature and immature part from \texttt{append} to \texttt{cons}.

After applying this two changes, stream representation becomes more complicated. It motivates us to set up an interface between stream functions and the rest of miniKanren. Listed in Fig.~\ref{interface} are all functions being aware of the stream representation, but \texttt{take-inf} and its helper function, which is explained later. The first three functions are constructors: \texttt{empty-inf} constructs an empty stream; \texttt{unit-mature-inf} constructs a stream with one mature solution; \texttt{unit-immature-inf} constructs a stream with one thunk. The \texttt{append-inf} in \texttt{mk-3} is relatively straightforwared compared with the \texttt{mk-1} version. \texttt{append-map-inf} is more tricky on how to construct the new immature part. We can follow the approach in mk-0 and mk-1 -- create a new thunk which invoke append-map-inf recursively when forced. But then we need to be careful: if we construct the thunk when the old immature part is an empty list, the resulting stream might be infinitely unproductive. Beside, all solutions of the next lowest cost in \texttt{s-inf} must be computed when the thunk is invoked. However sometimes only a portion of these solutions is required to answer a query. To avoid the trouble and the advanced computation, we choose to create a new thunk for every existing thunk. The next four functions are used only by \texttt{ifte} and \texttt{once}. Uninterested readers might skip them. \texttt{null-inf?} checks whether a stream is exausted. \texttt{mature-inf?} checks whether a stream has some mature solutions. \texttt{car-inf} takes the first solution out of a mature stream. \texttt{cdr-inf} drops the first solution of a mature stream. Finally, \texttt{force-inf} forces an a immature stream to do more computation.

\begin{figure}
	 \lstinputlisting{Figures/interface.rkt}	
	 \caption{Functions being aware of stream representation}
	 \label{interface}
\end{figure}

The last interesting function is \texttt{take-inf} (Fig.~\ref{take-inf-3-1}). The parameter \texttt{vs} is a list of solutions. The next two parameters together represents a functional queue in a typical way. The first two \texttt{cond} lines are very similar to their counterparts in mk-0 and mk-1. The third line runs when we exaust all solutions. The forth line re-shape the queue. The fifth and last line invoke the first thunk in the queue and use the mature part of the resulting stream, \texttt{s-inf}, as the new \texttt{vs}, and enqueuing \texttt{s-inf}'s thunks.

\begin{figure}
	 \lstinputlisting{Figures/take-inf-3-1.rkt}	
	 \caption{\texttt{take-inf} in \texttt{mk-3-1}}
	 \label{take-inf-3-1}
\end{figure}
